---
title: "VST Plot Mapper notes"
author: "Courtney Meier"
date: "6/23/2017"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(dplyr)
```

##  Adding a Leaflet map
### Resources
* https://rstudio.github.io/leaflet/shiny.html
* Leaflet tutorial using third-party tiles: https://www.youtube.com/watch?v=7UzD0P2BggM

### Layout of Shiny app
* 4 tabs:
    - Site Mapper: Leaflet map of all plots at selected site
    - Plot Mapper: ggplot of all mapped individuals within the selected plot from site selected on Site Mapper tab, option to download .pdf
    - Plot Data: Table of all tagged individuals within the selected plot, option to download .csv
    - About: Description of functionality for previous 3 tabs
* Site Mapper tab ideas:
    - Floating side panel on the left, main background panel for displaying map
    - Map: Build plots/subplots for map using plotSpatial and pointSpatial data, and SpatialPolygons Data Frame; already have code for this on devTOS for VST data checking.
    - Map: Initial zoom to North America?
    - Map: Auto-zoom to range of UTMs provided for plots at the site
    - Map: Clicking on a plot displays -> plotID, count of tagged stems
    - Map: Color of plots is based on color gradient associated with stem count (density)
    - Sidebar: Dropdown -> List of sites for which VST data are available
    - Sidebar: Dropdown -> Fill colors for plot polygons -> red, blue, orange, green
    - Sidebar: Radio buttons for map background -> Imagery (ESRI WorldMap), topographic (NatGeo?)
    - Sidebar: Control which variable is used to color plots: e.g., Density plot of stem count, based on plot/subplot data for all plots/subplots at selected site; color gradient fill using color selected in dropdown above; also could color plots by number of problem records within the plot.
    - Sidebar: Download screenshot of map? -> This may be difficult
    
    

### Dev ideas, next steps
* Code to identify duplicates expected for **tagID**, and select the latest value
```
si <- select(si, -uid)
si <- distinct(si)
si_last <- si %>%
            group_by(individualID, date, phenophaseName) %>%
            filter(editedDate==max(editedDate))
```

* Modify global.R to remove lat/long from `plotSpatial` -> going to use spatial data from `pointSpatial` for display instead -> need to figure out whether UTMs can be used with Leaflet, or whether decimal lat/long is needed.
* Begin testing of Leaflet code in normal R in environment, use ORNL as test case.


##  VegStructure Mapping Errors
Discoveries related to VegStructure data during development of Plot Mapper Shiny app:

* 10 records have no subplotID identified, need to have this field be required in future.
* Technicians are using additional pointIDs than those specified for mapping.

Quantifying mapping errors:

```{r pointErrors}
# pointids in vstInput dataset
usedPoints <- sort(unique(vstInput$pointid))

# pointids expected based on protocol
expPoints <- sort(unique(as.integer(pointSpatial$pointid)))
extraPoints <- setdiff(usedPoints, expPoints)

# Summary of rows in 'vstInput' associated with unexpected pointids
vstInput %>% filter(pointid %in% extraPoints) -> mapError
mapError %>% count(siteid, pointid)
mapError %>% count(siteid)
nrow(mapError)

### pointid errors
vstInput %>% 
  mutate(pointstatus = ifelse(is.na(pointid), "notMapped",
                              ifelse(plottype=="lgTower" & pointid %in% expLarge, "validPointID",
                                     ifelse(pointid %in% expSmall, "validPointID",
                                            "errorPointID")))) -> vstInput

```

### Summary of mapping errors

* Total number of mapped records affected = 720
* BART has greatest number of affected records = 541
* HARV has second most = 68
* D01 represents 85% of the problem records.

Closer examination of problems:

* `mapStatus == pointIDError` assigned if chosen point not one of expected points based on plottype (e.g., smTower, lgTower, distributed).
* BART pointIDError = 549, which is high as expected.
* SRER pointIDError = 228, which is much higher than expected.
* Code for plotSpatial table assigned Tower Plots == "smTower", based on plotsize=="400" being present in list of plotsize values for SRER Tower Plots -> is the plotSpatial table correct? I thought that SRER had large Tower Plots?
* At SRER, all pointIDError values associated with pointID = 23,39,43,59 which are perimeter cardinal direction pointIDs that would be acceptable for mapping in a lgTower plot, but would not be acceptable in a smTower plot. Result suggests that plotSize values for SRER Tower Plots in Plot Spatial data input file are incorrect.

### Making the ggplot

* Helpful hints for axes here: http://www.cookbook-r.com/Graphs/Axes_(ggplot2)/
* Using conditionals to add plot features - will be useful for drawing geom_segments() depending on subplotid: http://stackoverflow.com/questions/39592827/conditional-track-in-a-plot-r-ggplot2
* Adding tagID labels: http://docs.ggplot2.org/current/geom_text.html
* Adding caption https://www.r-bloggers.com/subtitles-and-captions-with-ggplot2-v-2-2-0/
* More interesting possibilities with text and text labels, this time with lines indicating which label goes to which dot: http://www.sthda.com/english/wiki/ggplot2-texts-add-text-annotations-to-a-graph-in-r-software
* Color mapped points by taxonID, make size dependent on diameter (latter eventually, since structure data aren't in Fulcrum yet)
* Consider making tick marks begin at origin (SW corner), and then going in 5 m increments from there; could also remove axis labels since techs may not use easting and northing data.
* For point labels on map, have 3 radio buttons: no labels, labels above points, and labels offset from points (repel function) with lines connected to point.
* Reactive function: Consider allowing user to change how mapped individuals are displayed (color coded, shape coded by taxonid) so that color blind techs can use the plot and techs can print on B&W paper.


### Next steps: Leaflet Map

* Blah